---
title: "Cutural factors of COViD-2019"
author: "Andrzej JArynowski"
date: "29/04/2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


## Problem Introducion
The COVID-19 outbreak data give an opportunity to comparare possible role of social attitudes and preferred social values.
Social norms are rules that prescribe what people should and should not do in a social surroundings. People are instructed by norms that tell them to keep their promises, or avoid deceiving other people. But how do norms work in a sudden change of social circumstances? 
Disregading medial differnces, we are trying to consider the relative cross-cultural differences to predict relative penetration of COVID-19 pandemic internationally.




#State of the art
There was few attempts to connect cultural variables:
- For 63 world countries, there was no significant correlation between cultural dimmnstions and COVID-19 cases (Pogrebna & Kharlamov 2020)
- For Poland, Italy, South Korea, China and Iran, there was visual realtionship cultural dimmnstions and success in COVID-19 menagment (Jarynowski, et al. 2020) 
- For 93 countries (Messner 2020) epidemic growth rate was explaned by institutional context (negatively associated) , education (positively associated),  individualistic rather than collectivistic (negatively associated),  higher levels of power distance (negatively associated), Hedonistic  values (positively associated).


```{r 1}
library(tidyverse)
library("ggrepel")

```

## Calculation of variables

We analyze cutural variables and time series of daily R0 estimates for 27 European countries.

We define staring R0 - maximal observed R0
Controlled R0 - mean R0 (Mean R0 for each country is a proxy of effectiveness of mitigation strategies)

Handwashing Avoidance = % of people who do not wash their hands after using toilet (BVA)
Cultural dimensions according to World Value Study (index in range 0-10):
"harmony"   
"embedded"     
"hierarchy"    
 "mastery"     
 "aff.auton"   
 "intel.auton"  
 "egalitar" 

```{r 2, echo=FALSE, include=TRUE}
cult_covid=read.csv("cultvalue_covid_eu.csv")
cult_covid$control=cult_covid$maxi/cult_covid$mediana
#colnames(cult_covid)

model=lm( control~harmony+embedded+hierarchy+mastery+aff.auton+intel.auton+egalitar+Handwashing , data = cult_covid)
#summary(model)
model_m=lm( maxi~harmony+embedded+hierarchy+mastery+aff.auton+intel.auton+egalitar+Handwashing , data = cult_covid)
#summary(model_m)
model_med=lm( mediana~harmony+embedded+hierarchy+mastery+aff.auton+intel.auton+egalitar+Handwashing , data = cult_covid)
#summary(model_med)

model_sred=lm( srednia~harmony+embedded+hierarchy+mastery+aff.auton+intel.auton+egalitar+Handwashing , data = cult_covid)
#summary(model_sred)

model_h=lm( control~harmony+embedded+hierarchy+mastery+aff.auton+intel.auton+egalitar, data = cult_covid)
#summary(model_h)

model_m=lm( maxi~harmony+embedded+hierarchy+mastery+aff.auton+intel.auton+egalitar , data = cult_covid)
#summary(model_m)
model_med=lm( mediana~harmony+embedded+hierarchy+mastery+aff.auton+intel.auton+egalitar , data = cult_covid)
#summary(model_med)
cult_covid$controlled_R0=cult_covid$srednia
cult_covid$starting_R0=cult_covid$maxi

```

## Results

We can see significant relationsip in a linear multiple regression between controlled R0 and cultural dimensions.

```{r tabela}
model_controlled=lm(controlled_R0~harmony+embedded+hierarchy+mastery+aff.auton+intel.auton+egalitar, data = cult_covid)
summary(model_controlled)

```
embedded and aff.auton are significantly negatively associated with control.   

hierarchy and intel.auton are significantly positively associated with control.

Remark - lower controlled R0 is better


Visualization of 2 most significant cultural dimensions an controlled/starting R0.
```{r wumiary}

cult_covid$controlled_Rt=cult_covid$controlled_R0
cult_covid$starting_Rt=cult_covid$starting_R0

pl <- ggplot(cult_covid, aes(embedded, aff.auton, size=starting_Rt, color=controlled_Rt)) + theme_bw()+
  geom_point() + xlab("embedded") +
geom_text_repel(aes(label=country_code),hjust=0, vjust=0, color="red")
pl

```

Hand hygene and starting R0.

```{r higiena}


pl <- ggplot(cult_covid, aes(Handwashing, starting_R0, size=starting_R0, color=controlled_R0)) +
  geom_point() + xlab("handwashing avoidance") +geom_text(aes(label=country_code),hjust=0, vjust=0, color="red")+ geom_smooth(method = "lm")
pl

```

## Condlusions

The results emphasize the need to pay more  attention to cultural context when compering outbreak dynamics between countries

Literature

JARYNOWSKI, A., WÓJTA-KEMPA, M., PŁATEK, D., & CZOPEK, K. (2020). ATTEMPT TO UNDERSTAND PUBLIC-HEALTH RELEVANT SOCIAL DIMENSIONS OF COVID-19 OUTBREAK IN POLAND. Society Register, 4(3), 7-44.
Pogrebna, G., & Kharlamov, A. (2020). The Impact of Cross-Cultural Differences in Handwashing Patterns on the COVID-19 Outbreak Magnitude.
Messner, W. (2020). The Institutional and Cultural Context of Cross-National Variation in COVID-19 Outbreaks. medRxiv.
